### 2020-04-21 Time_Series_Analysis.Rmd
Create Time_Series_Analysis.Rmd
time series analysis of China

---
title: "COVID-19 Forecasting"
author: "Ashar Kandathil"
date: "21/04/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(ggplot2)
library(zoo)
library(tseries)
library(forecast)
```

## Intro

We hope to create a time series analysis of the global COVID-19 situation. Using ARIMA modeling, we wish to generate useful forecasts of what the situation could be like before and after various quarantine and social-distancing protocols were enacted, to see whether the "curve is flattening" and to what extent. For this report analyses we will only focus on 1 country. 
For the rest of the countries we will forecast the further spread given the latest data, to be displayed on our Rshiny dashboard.
We understand that this is an extremely complex topic with numerous factors that determine the rate of spread, death, and recoveries for any given country. However, to simplify things we will carry on with the assumption that in order to combat the spread, there was a specific date in which preventative measures were enabled for the given country.

Note: Our global data starts from January 22, 2020. The first cases of the virus were recorded in November of 2019. Given our small dataset, we assume non-seasonality.
```{r cars}
# READ AND AUGMENT DATA

covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
dates <- unique(covid_data$Date)
firstDate = dates[1] # Jan 22, 2020

confirmed_df <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
death_df <- aggregate(Deaths ~ Country.Region + Date, FUN = sum, data=covid_data)
recovery_df <- aggregate(Recovered ~ Country.Region + Date, FUN = sum, data=covid_data)

partial_merge <- merge(confirmed_df, death_df, by = c("Country.Region", "Date"))
covid_df <- merge(partial_merge, recovery_df, by = c("Country.Region", "Date"))

countries <- unique(covid_df$Country.Region)
countries_ts <- lapply(countries, function (x) {
  country_ts <- covid_df %>% filter(Country.Region == x) %>% select(c(Confirmed, Deaths, Recovered))
  ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
})

```

```{r pressure, echo=FALSE}
# FUNCTION DEFINITIONS

get_country_ts <- function(country_name){
  idx = which(countries == country_name)[[1]]
  countries_ts[[idx]]
}

augmented_dickey_fuller <- function(c_ts, series, regular, lag_order) {
  if(regular==FALSE){
    adf.test(diff(log(c_ts[, series])), alternative = "stationary", k = lag_order)
  } else {
    adf.test(c_ts[, series], alternative = "stationary", k = lag_order)
  }
}

date_filtered_zoo <- function(country_name, endDate) {
  wantedDates <- dates[between(dates, firstDate, as.Date(endDate))]
  country_ts <- covid_df %>% filter(Country.Region == country_name, between(dates, firstDate, as.Date(endDate))) %>% select(Confirmed)
  zoo(country_ts, wantedDates)
}

plot_zoo <- function(country_zoo, country_name) {
  autoplot(country_zoo, main = paste("COVID-19 Confirmed Cases in", country_name)) + labs(x = "Date", y = "Number of People")
}

plot_ts <- function(country_ts, country_name) {
  autoplot(country_ts, main = paste("COVID-19 Statistics in", country_name, "Since Jan 22")) + labs(x = "Time", y = "Number of People") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
}
```

## China

Let us begin by looking at the country where the disease was first encountered - China. By looking into how the disease spread and China handled the situation first, we may be able anticipate how other countries should handle the situation.

```{r}
# China
china_zoo <- date_filtered_zoo("China", "2020-4-12")
plot_zoo(china_zoo, "China")
```
The first graph we want to look at is the number of confirmed cases in China to date. As we can see from our preliminary plot, there was a sharp spike in cases from January till March, after which the curve appears to be "flattened".

From external research, it appears that China began its lockdown measures on Wuhan around January 29, which continued to rollout across the country. (https://www.edinburghnews.scotsman.com/health/how-long-has-china-been-lockdown-heres-when-china-and-wuhan-went-lockdown-coronavirus-restrictions-approach-end-2514773). However, it takes a couple weeks for the effects of such measures to show, as people can be asymptomatic for 14 days. For simplicity, we will assume that February 12 (2 weeks after preventative measures) as the cut-off point for all of China.

Let us forecast the number of potential cases that would exist in China today, if no preventative measures were taken. 
### Analysis
```{r}
before_china_zoo <- date_filtered_zoo("China", "2020-2-12")
plot_zoo(before_china_zoo, "China")
```
To start with, we must first determine the stationarity of our time series with the Augmented Dickey-Fuller test. We can assume a k-value (lag) of *n/5 = 4.2 ~ 4* as recommended for non-seasonal series [Hyndman].
```{r}
china_ts <- get_country_ts("China")
before_china_ts <- head(china_ts, 21) # Feb 12 is 3 weeks from Jan 22 = 21 days
augmented_dickey_fuller(before_china_ts, "Confirmed", TRUE, 4)
```
According to the test, it appears that the data is non-stationary given the high p-value. Let us remove the unequal variances and trend components, by taking the log and difference of the series, respectively.
```{r}
augmented_dickey_fuller(before_china_ts, "Confirmed", FALSE, 4)
```
Now the p-value is very small and clearly significant. Hence, the data is now stationary.

Let us plot the ACF and PACF charts to determine the *p* and *q* portions of our ARIMA model.
```{r}
ggAcf(diff(log(before_china_ts[,"Confirmed"]))) + ggtitle("Confirmed Cases ACF Plot")
ggPacf(diff(log(before_china_ts[,"Confirmed"]))) + ggtitle("Confirmed Cases PACF Plot")
```
It appears that the series is mostly represented by an Auto-Regressive (AR) model due to the entire PACF plot cutting off and the ACF plot only cutting off after the 2nd lag. We would assume an ARIMA(0,1,2) model. However, it is wise to experiment with different combinations to find the best ARIMA(p,d,q) values that have the lowest AIC and BIC.

Rather than conduct this manually, we will use the `auto.arima()` function to calculate it automatically for us, with the `seasonal`, `stepwise`, and `approximation` parameters set to `FALSE` so that the best model can be chosen.
```{r}
china_fit <- auto.arima(before_china_ts[, "Confirmed"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
china_fit
```
As we see from the output, our series follows an ARIMA(1,2,0) model (contrary to our initial estimation of ARIMA(0,1,2)).

Let us now forecast the infection rate if China had not taken any preventative measures, for a month after February 12.
```{r}
china_fore <- forecast(china_fit, h = 30, level = 95)
plot(china_fore, ylab = "Number of People", main = "Forecasts for Confirmed Cases from Feb 12", xlab = "Time", xaxt = "n")
```
As seen from the line in blue, our forecasting till March 12 is approximately 100,000 infections (20,000 higher than the Confirmed cases by April 12), with 95% *prediction intervals* going upwards of 200,000 infections.

Therefore, we can infer that the preventative measure to slow the spread of COVID-19 truly are "flattening the curve".

### 2020-04-21 test.Rmd
added functions

---
title: "A test ting"
output: html_notebook
---

```{r}
library(ggplot2)
library(plyr)
library(dplyr)
library(gridExtra)
library(GGally) #for ggpairs
library(psych)
library(ggiraph)
library(ggiraphExtra)
```

```{r}
# IMPORTANT must import the data and apply these manipulations first. Otherwise the function will not work

covid_set <- read.csv('data/covid_data/covid_19_clean_complete.csv')
covid_set <- covid_set %>% mutate(date_str =  as.Date(as.POSIXct(Date, format="%m/%d/%y"))) %>% arrange(date_str)
covid_set

```

```{r}
# plotting deaths function
# USAGE: plot_deaths('Canada')
plot_deaths <- function(country) {
  country_set <- covid_set %>% filter(Country.Region == country)
  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
  deaths <- ggplot(country_set.agg, mapping = aes(date_str, Deaths)) + geom_line() + labs(title='Deaths', x='Date', y='Deaths')
  return(deaths)
}

# plotting confirmed function
# USAGE: plot_confirmed('Canada')
plot_confirmed <- function(country) {
  country_set <- covid_set %>% filter(Country.Region == country)
  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
  confirmed <- ggplot(country_set.agg, mapping = aes(date_str, Confirmed)) + geom_line() + labs(title='Confirmed Cases', x='Date', y='Cases')
  return(confirmed)
}

# plotting recovered function
# USAGE: plot_confirmed('Canada')
plot_recovered <- function(country) {
  country_set <- covid_set %>% filter(Country.Region == country)
  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
  recovered <- ggplot(country_set.agg, mapping = aes(date_str, Recovered)) + geom_line() + labs(title='Recovered', x='Date', y='Recovered')
  return(recovered)
}

plot_deaths('Canada')
plot_confirmed('Canada')
plot_recovered('Canada')
```

```{r}
# plotting change in deaths
# USAGE plot_change_deaths('Canada')
plot_change_deaths <- function(country) {
  country_set <- covid_set %>% filter(Country.Region == country)
  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
  country_set.new_by_week <- country_set.agg %>% mutate(new_deaths_by_week=Deaths - lag(Deaths, 7))
  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title='Trajectory of Deaths', x='New Deaths (past week)', y='Total Deaths')
  return(plot)
}

# plotting change in cases
# USAGE plot_change_confirmed('Canada')
plot_change_confirmed <- function(country) {
  country_set <- covid_set %>% filter(Country.Region == country)
  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
  country_set.new_by_week <- country_set.agg %>% mutate(new_confirmed_by_week=Confirmed - lag(Confirmed, 7))
  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title='Trajectory of Confirmed Cases', x='New Confirmed Cases (past week)', y='Total Confirmed Cases')
  return(plot)
}

plot_change_deaths('Canada')
plot_change_confirmed('Canada')
```

### 2020-04-21 test.Rmd
labelled axis fix

@@
   country_set.new_by_week <- country_set.agg %>% mutate(new_deaths_by_week=Deaths - lag(Deaths, 7))
-  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title='Trajectory of Deaths', x='New Deaths (past week)', y='Total Deaths')
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title='Trajectory of Deaths', y='New Deaths (past week)', x='Total Deaths')
   return(plot)
@@
   country_set.new_by_week <- country_set.agg %>% mutate(new_confirmed_by_week=Confirmed - lag(Confirmed, 7))
-  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title='Trajectory of Confirmed Cases', x='New Confirmed Cases (past week)', y='Total Confirmed Cases')
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title='Trajectory of Confirmed Cases', y='New Confirmed Cases (past week)', x='Total Confirmed Cases')
   return(plot)

### 2020-04-21 Time_Series_Analysis.Rmd
Update Time_Series_Analysis.Rmd

@@
 
-Note: Our global data starts from January 22, 2020. The first cases of the virus were recorded in November of 2019. Given our small dataset, we assume non-seasonality.
-```{r cars}
+Note: Our global data starts from January 22, 2020. However, in reality the first cases of the virus were recorded in November of 2019. Given our small dataset, we assume non-seasonality.
+```{r, include = FALSE}
 # READ AND AUGMENT DATA
@@
 
-```{r pressure, echo=FALSE}
+```{r, include=FALSE}
 # FUNCTION DEFINITIONS
@@
 
-## China
-
 Let us begin by looking at the country where the disease was first encountered - China.
@@
 
-```{r}
+```{r, echo=FALSE}
 # China
@@
 
-From external research, it appears that China began its lockdown measures on Wuhan around January 29, which continued to rollout across the country.
-(https://www.edinburghnews.scotsman.com/health/how-long-has-china-been-lockdown-heres-when-china-and-wuhan-went-lockdown-coronavirus-restrictions-approach-end-2514773).
+From external research, it appears that China began its lockdown measures on Wuhan around January 29, which continued to rollout across the country [Greig].
 However, it takes a couple weeks for the effects of such measures to show, as people can be asymptomatic for 14 days.
@@
 
-Let us forecast the number of potential cases that would exist in China today, if no preventative measures were taken. 
-### Analysis
-```{r}
+Let us forecast the number of potential cases that would exist in China if no preventative measures were taken.
+
+## Analysis
+```{r, echo=FALSE}
 before_china_zoo <- date_filtered_zoo("China", "2020-2-12")
@@
 We can assume a k-value (lag) of *n/5 = 4.2 ~ 4* as recommended for non-seasonal series [Hyndman].
-```{r}
+```{r, echo=FALSE}
 china_ts <- get_country_ts("China")
@@
 Let us remove the unequal variances and trend components, by taking the log and difference of the series, respectively.
-```{r}
+```{r, echo=FALSE}
 augmented_dickey_fuller(before_china_ts, "Confirmed", FALSE, 4)
@@
 Let us plot the ACF and PACF charts to determine the *p* and *q* portions of our ARIMA model.
-```{r}
+```{r, echo=FALSE}
 ggAcf(diff(log(before_china_ts[,"Confirmed"]))) + ggtitle("Confirmed Cases ACF Plot")
@@
 Rather than conduct this manually, we will use the `auto.arima()` function to calculate it automatically for us, with the `seasonal`, `stepwise`, and `approximation` parameters set to `FALSE` so that the best model can be chosen.
-```{r}
+```{r, echo=FALSE}
 china_fit <- auto.arima(before_china_ts[, "Confirmed"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
@@
 Let us now forecast the infection rate if China had not taken any preventative measures, for a month after February 12.
-```{r}
-china_fore <- forecast(china_fit, h = 30, level = 95)
+```{r, echo=FALSE}
+china_fore <- forecast(china_fit, h = 30, level = 90)
 plot(china_fore, ylab = "Number of People", main = "Forecasts for Confirmed Cases from Feb 12", xlab = "Time", xaxt = "n")
 ```
-As seen from the line in blue, our forecasting till March 12 is approximately 100,000 infections (20,000 higher than the Confirmed cases by April 12), with 95% *prediction intervals* going upwards of 200,000 infections.
+As seen from the line in blue, our forecasting till March 12 is approximately 100,000 infections (20,000 higher than the Confirmed cases by April 12), with 90% *prediction intervals* going upwards of 200,000 infections.
 
 Therefore, we can infer that the preventative measure to slow the spread of COVID-19 truly are "flattening the curve".
+
+## References
+
+Greig, Finlay.
+“How Long Has China Been on Lockdown?
+Here's When China and Wuhan Went on Lockdown for Coronavirus.” Edinburgh News, Edinburgh News, 6 Apr.
+2020, www.edinburghnews.scotsman.com/health/how-long-has-china-been-lockdown-heres-when-china-and-wuhan-went-lockdown-coronavirus-restrictions-approach-end-2514773.
+
+Hyndman, Robert. “Thoughts on the Ljung-Box Test.” Hyndsight, 24 Jan. 2014, www.robjhyndman.com/hyndsight/ljung-box-test/.

### 2020-04-21 ts_forecast_function.R
forecast function

@@ -1,0 +1,13 @@
+covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
+covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
+covid_data <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
+
+forecast_confirmed <- function(country) {
+  country_ts <- covid_data %>% filter(Country.Region == country) %>% select(c(Confirmed))
+  country_ts <- ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
+  fit <- auto.arima(country_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
+  fore <- forecast(fit, h = 30, level = 90)
+  plot(fore, ylab = "Number of People", main = paste("Forecast for Number of Cases in", country), xlab = "Time", xaxt = "n")
+}
+
+forecast_confirmed("Canada")

### 2020-04-22 app.R
implemented covid map and graphs

@@ -5,47 +5,152 @@ library(nominatim)
 library(tmap)
 library(tidygeocoder)
 library(dplyr)
+library(shinythemes)
+library(ggplot2)
+library(shinyjs)
 
+### Importing Data ###
 country_df <- read.csv('data/country_data/countries of the world.csv')
 coordinate_df <- read.csv('data/country_data/countries.csv')
+covid_df <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+covid_set <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+covid_set <- covid_set %>% mutate(date_str =  as.Date(as.POSIXct(Date, format="%m/%d/%y"))) %>% arrange(date_str)
+covid_set
+
+### Cleaning and formatting Data ###
+
+# removes trailing whitespaces
 country_df[,c(1)]=trimws(country_df[,c(1)])
+
+# renaming columns to conform to the leaflet lat / long naming convention
 names(coordinate_df)[names(coordinate_df) == "name"] <- "Country"
 names(coordinate_df)[names(coordinate_df) == "latitude"] <- "lat"
 names(coordinate_df)[names(coordinate_df) == "longitude"] <- "long"
+# joining lat long dictionary to our high level country data
 mergedData <<- merge(x=country_df, y=coordinate_df, by="Country", all.x=TRUE)
+# drops rows with any NA in them
 cleanData <<- mergedData[complete.cases(mergedData), ]
-print(cleanData)
+# drops rows with any empty in them
+cleanData[!apply(cleanData == "", 1, all),]
+# get a vector of names of cols for dimension input
+display_vec <<- unique(names(cleanData))
+display_vec <<- display_vec[display_vec != c('Country', 'Region')]
+# For some reson only a staggred removal is rendering in the ui ...
+display_vec <<- display_vec[display_vec != c('country', 'lat', 'long')]
 
+# test to join covid and geo data
+names(covid_df)[names(covid_df) == "Country.Region"] <- "Country"
+names(covid_df)[names(covid_df) == "Lat"] <- "lat"
+names(covid_df)[names(covid_df) == "Long"] <- "long"
+main_df <<- merge(x=covid_df, y=country_df, by="Country", all.x=TRUE)
 
-ui <- bootstrapPage(
+### Helper functions ###
+
+# Transparent theme enforced here (apparentlty you have to double enfore it in renderOutput as well)
+# TODO - see how to make semi transparent ...
+transparent_theme <-  theme(
+  panel.grid.major = element_blank(),
[…]

### 2020-04-22 MSCI718 2020W Project v3.Rmd
started report

---
title: "MSCI 718 Mini-Project"
author: "Matthieu Court, Ashar Kandathil, Tyler Lam, Kevin Lee"
date: "April 22, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Instructions: Please complete this assignment and submit online to the LEARN Dropbox as a PDF; the PDF will be considered complete for marking. Also upload any source files, for example, any .R or .Rmd files that you used.*

The final project deliverable must include a **PDF report (max 6 pages, references not included)** and any source code for your project.

**NOTE:** the emphasis **must** be on the application, interpretation, and communication of the statistical method. As part of this, you are **not allowed** to do black-box methods like neural nets. As you've seen, this course is on working data in a transparent and interpretable manner, not on big data analytics.

#Data
Due to the current state of the world from the COVID-19 pandemic, this report focused on using exploratory data analysis, time series forecasting, and data visualization to forecast the number of incidents, recoveries and deaths on a per country basis.

The Global Coronavirus Dataset was sourced from Kaggle and used for this analysis. 
There were three appropriate datasets that were sourced from Kaggle:
1. Global Coronavirus Dataset

We found a dataset that contains information on every country’s number of confirmed cases, deaths, and recoveries since January 1, 2020.

2. Countries of the World Fact Sheet Dataset
3. UN Country Statistics Dataset

We also found 2 other datasets that contain national information on each country, ranging from population to GDP per capita. By combining these 2 datasets, we want to create an interactive dashboard that forecasts the number of cases, recoveries, and deaths per country based on potential correlation with other factors.

Datasets:

#Analysis

#Conclusion

### 2020-04-22 app.R
wip

@@ -8,14 +8,22 @@ library(dplyr)
 library(shinythemes)
 library(ggplot2)
 library(shinyjs)
+library(zoo)
+library(tseries)
+library(forecast)
 
 ### Importing Data ###
 country_df <- read.csv('data/country_data/countries of the world.csv')
 coordinate_df <- read.csv('data/country_data/countries.csv')
 covid_df <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+# Prediction import
 covid_set <- read.csv('data/covid_data/covid_19_clean_complete.csv')
 covid_set <- covid_set %>% mutate(date_str =  as.Date(as.POSIXct(Date, format="%m/%d/%y"))) %>% arrange(date_str)
 covid_set
+# Time series import
+covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
+covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
+covid_data <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
 
 ### Cleaning and formatting Data ###
 
@@ -97,22 +105,46 @@ plot_change_confirmed <- function(country) {
   return(plot)
 }
 
+forecast_confirmed <- function(country) {
+  country_ts <- covid_data %>% filter(Country.Region == country) %>% select(c(Confirmed))
+  country_ts <- ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
+  fit <- auto.arima(country_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
+  fore <- forecast(fit, h = 30, level = 90)
+  graphics.off()
+  #colnames(fore) <-  c("time", "point_forecast", "lo_90", "hi_90")
+  print(fore)
+  
+  # Create a dataframe from 'fore'
+  result_df <- as.data.frame(fore)
+  result_df <- tibble::rownames_to_column(result_df, "point")
+  rownames(result_df) <- NULL
+  colnames(result_df) <- c("point","forecast","lo", "hi")
+  print(result_df)
+
+  tester <- plot(result_df$point, result_df$forecast)
+  tester_2 <- ggplot(data = result_df, mapping = aes(x = point, y = forecast)) + geom_line()
[…]

### 2020-04-22 ts_forecast_function.R
Update ts_forecast_function.R
return plot as ggplot @mattlianje

@@ -6,8 +6,11 @@ forecast_confirmed <- function(country) {
   country_ts <- covid_data %>% filter(Country.Region == country) %>% select(c(Confirmed))
   country_ts <- ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
   fit <- auto.arima(country_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
-  fore <- forecast(fit, h = 30, level = 90)
-  plot(fore, ylab = "Number of People", main = paste("Forecast for Number of Cases in", country), xlab = "Time", xaxt = "n")
+  fore <- forecast(fit, h = 30, level = 90) 
+  autoplot(fore, ylab = "Number of People", main = paste("Forecast for Number of Cases in", country), xlab = "Time", xaxt = "n") +
+    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
 }
 
 forecast_confirmed("Canada")
+
+

### 2020-04-22 app.R
added ts projection, fixed ui, got latest csv

@@ -62,6 +62,12 @@ transparent_theme <-  theme(
   panel.background = element_rect(fill = "transparent",colour = NA),
   plot.background = element_rect(fill = "transparent",colour = NA)
 )
+
+white_theme <- theme(
+  panel.grid.major = element_blank(),
+  panel.grid.minor = element_blank(),
+  plot.background = element_rect(fill = "transparent",colour = NA)
+)
 # plotting deaths function
 # USAGE: plot_deaths('Canada')
 plot_deaths <- function(country) {
@@ -110,8 +116,6 @@ forecast_confirmed <- function(country) {
   country_ts <- ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
   fit <- auto.arima(country_ts, seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
   fore <- forecast(fit, h = 30, level = 90)
-  graphics.off()
-  #colnames(fore) <-  c("time", "point_forecast", "lo_90", "hi_90")
   print(fore)
   
   # Create a dataframe from 'fore'
@@ -119,12 +123,10 @@ forecast_confirmed <- function(country) {
   result_df <- tibble::rownames_to_column(result_df, "point")
   rownames(result_df) <- NULL
   colnames(result_df) <- c("point","forecast","lo", "hi")
-  print(result_df)
-
-  tester <- plot(result_df$point, result_df$forecast)
-  tester_2 <- ggplot(data = result_df, mapping = aes(x = point, y = forecast)) + geom_line()
   final_plot <- plot(fore, ylab = "Number of People", main = paste("Forecast for Number of Cases in", country), xlab = "Time", xaxt = "n")
-  return(final_plot)
+  ashar_plot <- autoplot(fore, ylab = "Number of People", main = paste("Forecast for Number of Cases in", country), xlab = "Time", xaxt = "n") +
+    theme_classic() + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+  return(ashar_plot)
 }
 
 forecast_confirmed("Canada")
@@ -134,17 +136,17 @@ ui <- bootstrapPage(theme = shinytheme('cosmo'),
   leafletOutput("map", width = "100%", height = "100%"),
   absolutePanel(top = 10, right = 20,
                 useShinyjs(),
-                sliderInput("date_slider", "Date (cumulative)", min = as.Date('2020-1-20'),
+                sliderInput("date_slider", "Date (cumulative)", min = as.Date('2020-1-22'),
[…]

### 2020-04-22 MSCI718 2020W Project v3.Rmd
Update MSCI718 2020W Project v3.Rmd

@@
 ---
-title: "MSCI 718 Mini-Project"
+title: "MSCI 718 Mini-Project: COVID-19 Time Series Analysis and Visualization"
 author: "Matthieu Court, Ashar Kandathil, Tyler Lam, Kevin Lee"
@@
 knitr::opts_chunk$set(echo = TRUE)
+library(lubridate)
+library(dplyr)
+library(ggplot2)
+library(zoo)
+library(tseries)
+library(forecast)
+library(gridExtra)
 ```
 
-*Instructions: Please complete this assignment and submit online to the LEARN Dropbox as a PDF; the PDF will be considered complete for marking.
-Also upload any source files, for example, any .R or .Rmd files that you used.*
+# Data
 
-The final project deliverable must include a **PDF report (max 6 pages, references not included)** and any source code for your project.
+Due to the current state of the world from the COVID-19 pandemic (as of April 22nd, 2020) and the rise of visualization tools being used to keep track of COVID-19 statistics across the world, we were interested in using concepts that were learned in MSCI 718 and new statistical techniques, more specifically time series analysis and interactive data visualiztion dashboard, to build a deeper understanding of the coronavirus.
 
-**NOTE:** the emphasis **must** be on the application, interpretation, and communication of the statistical method.
-As part of this, you are **not allowed** to do black-box methods like neural nets.
-As you've seen, this course is on working data in a transparent and interpretable manner, not on big data analytics.
+Firstly, a Global COVID-19 Report dataset was sourced from Kaggle with 15,170 observations.
+This dataset included daily entries of cumulative number of cases of COVID-19 for 185 countries, each with 82 entries dating from January 22nd, 2020 to April 12th, 2020.
+The following were the relevant variables from the dataset that were used for the analysis:
 
-#Data
-Due to the current state of the world from the COVID-19 pandemic, this report focused on using exploratory data analysis, time series forecasting, and data visualization to forecast the number of incidents, recoveries and deaths on a per country basis.
+* Country/Region: country of the observation
+* Date: date of the observation
+* Confirmed: cumulative number of confirmed cases on this date
 
-The Global Coronavirus Dataset was sourced from Kaggle and used for this analysis. 
-There were three appropriate datasets that were sourced from Kaggle:
-1. Global Coronavirus Dataset
+```{r,echo=FALSE,include=FALSE}
+covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
+covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
+dates <- unique(covid_data$Date)
 
-We found a dataset that contains information on every country’s number of confirmed cases, deaths, and recoveries since January 1, 2020.
+firstDate = dates[1] # Jan 22, 2020
+confirmed_df <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
+death_df <- aggregate(Deaths ~ Country.Region + Date, FUN = sum, data=covid_data)
+recovery_df <- aggregate(Recovered ~ Country.Region + Date, FUN = sum, data=covid_data)
+partial_merge <- merge(confirmed_df, death_df, by = c("Country.Region", "Date"))
+covid_df <- merge(partial_merge, recovery_df, by = c("Country.Region", "Date"))
+countries <- unique(covid_df$Country.Region)
+countries_ts <- lapply(countries, function (x) {
+  country_ts <- covid_df %>% filter(Country.Region == x) %>% select(c(Confirmed, Deaths, Recovered))
+  ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
+})
 
-2. Countries of the World Fact Sheet Dataset
-3. UN Country Statistics Dataset
+summary(confirmed_df)
+summary(confirmed_df %>% group_by(Country.Region) %>% summarise(count = n()))
+```
+The Confirmed variable ranged from 0 cases reported all the way to 555,313 cases reported, with the mean being 1,800 and median being 0.
+The central tendency of the data doesn't tell us much since these were cumulative numbers, but it could be understood that more than half of the reported numbers were 0, with an exponential increase as the date progressed.
+The outliers of cases on the right of the frequency graph are from United States, which have more confirmed cases in the world than any other country by a large margin.
+However, due to the approach of the analysis, the outliers would not affect the results of the models.
+Therefore, no treatment was required.
 
-We also found 2 other datasets that contain national information on each country, ranging from population to GDP per capita.
-By combining these 2 datasets, we want to create an interactive dashboard that forecasts the number of cases, recoveries, and deaths per country based on potential correlation with other factors.
+The frequency distribution of confirmed cases can be viewed in the histogram below.
 
-Datasets:
+```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.8}
+a <- ggplot(confirmed_df,aes(Confirmed)) + geom_histogram() + stat_bin(aes(y=..count.., label =..count..),geom="text",size=2,vjust=-0.3) + ggtitle("Frequency of Confirmed Cases") + ylab("Frequency")
+a
+```
+&nbsp;
 
-#Analysis
+In the Cumulative Confirmed Cases Over Time plot below, it can be observed that COVID-19 cases have been exponentially increasing in many countries across the world.
+This trend will be more thoroughly analyzed in the time series analysis.
 
-#Conclusion
+```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=3}
+
+b <- ggplot(confirmed_df, aes(Date, Confirmed,group=Country.Region)) + geom_point(size = 0.5) + geom_line() + ylab("Cumulative Confirmed Cases") + ggtitle("Cumulative Confirmed Cases Over Time")
+b
+```
+
+# Analysis
+
+# Conclusion

### 2020-04-23 app.R
removed hardocded dates in slider

@@ -51,6 +51,7 @@ names(covid_df)[names(covid_df) == "Country.Region"] <- "Country"
 names(covid_df)[names(covid_df) == "Lat"] <- "lat"
 names(covid_df)[names(covid_df) == "Long"] <- "long"
 main_df <<- merge(x=covid_df, y=country_df, by="Country", all.x=TRUE)
+main_df$Date <- as.Date(main_df$Date, format="%m/%d/%y")
 
 ### Helper functions ###
 
@@ -136,8 +137,8 @@ ui <- bootstrapPage(theme = shinytheme('cosmo'),
   leafletOutput("map", width = "100%", height = "100%"),
   absolutePanel(top = 10, right = 20,
                 useShinyjs(),
-                sliderInput("date_slider", "Date (cumulative)", min = as.Date('2020-1-22'),
-                            max =as.Date('2020-5-20'),value=as.Date("2020-4-01"),timeFormat="%b %d - %Y"),
+                sliderInput("date_slider", "Date (cumulative)", min = min(main_df$Date),
+                            max = max(main_df$Date),value=as.Date("2020-4-01"),timeFormat="%b %d - %Y"),
                 selectInput("dimension", "Country (projections)",
                             main_df$Country, selected = "Canada", multiple = FALSE
                 ),
@@ -147,7 +148,7 @@ ui <- bootstrapPage(theme = shinytheme('cosmo'),
                 plotOutput("plot_change_confirmed", height="125px"),
                 plotOutput("plot_deaths", height="125px"),
                 plotOutput("time_series", height="125px"),#,
-                #checkboxInput("legend", "Show legend", TRUE)
+                #checkboxInput("legend", "Show legend", TRUE)as.Date(as.yearmon(dataset$minDate))
   )
 )
 

### 2020-04-23 test.Rmd

@@
   country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
-  deaths <- ggplot(country_set.agg, mapping = aes(date_str, Deaths)) + geom_line() + labs(title='Deaths', x='Date', y='Deaths')
+  deaths <- ggplot(country_set.agg, mapping = aes(date_str, Deaths)) + geom_line() + labs(title=paste('Deaths', country), x='Date', y='Deaths')
   return(deaths)
@@
   country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
-  confirmed <- ggplot(country_set.agg, mapping = aes(date_str, Confirmed)) + geom_line() + labs(title='Confirmed Cases', x='Date', y='Cases')
+  confirmed <- ggplot(country_set.agg, mapping = aes(date_str, Confirmed)) + geom_line() + labs(title=paste('Confirmed Cases ', country), x='Date', y='Cases')
   return(confirmed)
@@
   country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
-  recovered <- ggplot(country_set.agg, mapping = aes(date_str, Recovered)) + geom_line() + labs(title='Recovered', x='Date', y='Recovered')
+  recovered <- ggplot(country_set.agg, mapping = aes(date_str, Recovered)) + geom_line() + labs(title=paste('Recovered ', country), x='Date', y='Recovered')
   return(recovered)
@@
   country_set.new_by_week <- country_set.agg %>% mutate(new_deaths_by_week=Deaths - lag(Deaths, 7))
-  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title='Trajectory of Deaths', y='New Deaths (past week)', x='Total Deaths')
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title=paste('Trajectory of Deaths', country), y='New Deaths (past week)', x='Total Deaths')
   return(plot)
@@
   country_set.new_by_week <- country_set.agg %>% mutate(new_confirmed_by_week=Confirmed - lag(Confirmed, 7))
-  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title='Trajectory of Confirmed Cases', y='New Confirmed Cases (past week)', x='Total Confirmed Cases')
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title=paste('Trajectory of Confirmed Cases', country), y='New Confirmed Cases (past week)', x='Total Confirmed Cases')
   return(plot)
@@
 ```
+
+# Are we Flattening the Curve?
+
+Governments all over the world have enacted social distancing measures in attempts to "flatten the curve".
+In more mathematical terms, flattening the curve means slowing the growth rate of the virus such that it is no longer on an exponential trajectory.
+This prevents the health care system from being overwhelmed and increase everyone's chances of surviving the virus.
+When we plot the confirmed cases over time it's difficult to visually tell if the growth is still exponential or if social distancing has slowed the growth.
+
+```{r echo=FALSE}
+require(gridExtra)
+
+p1 <- plot_confirmed('South Korea') + theme(plot.title = element_text(size = 10))
+p2 <- plot_change_confirmed('South Korea') + theme(plot.title = element_text(size = 10))
+
+y <- c(0,2,4,8,16,32,64,128,256)
+x <- c(0,2,6,14,30,62,126,254,510)
+p3 <- ggplot(data.frame(x,y), aes(x, y)) + geom_line() + geom_point() + labs(title='y=2^x', x='Total Confirmed Cases', y='New Cases (past week)') + theme(plot.title = element_text(size = 10))
+
+grid.arrange(p1, p2, p3, ncol=3)
+```
+
+A trajectory graph better illustrates the state of flattening the curve.
+The y axis represents the new confirmed cases in the past week, and the x axis represents the total confirmed cases.
+If the growth rate is exponential we can expect the graph to have a positive linear pattern.
+This can be demonstrated by graphing the function y=2^x.
+Once the growth has slowed and is no longer exponential, there a sharp dip in the graph.
+
+A country that has done really well in containing the virus is South Korea.
+As demonstrated by the linear growth, South Korea started with an exponential growth like every other country, but their effective social distancing and rapid testing measures is reflected in the graph's sharp dip.
+
+Using this graph provides a clear visual indication if a country has beaten the exponential growth and is on the right track towards flattening the curve.

### 2020-04-23 app.R
implemented covid map and graphs (#5)
* implemented covid map and graphs

* wip

* added ts projection, fixed ui, got latest csv

* removed hardocded dates in slider

Co-authored-by: Matthieu Court <mcourt@Matthieus-MacBook-Air.local>

@@ -5,47 +5,201 @@ library(nominatim)
 library(tmap)
 library(tidygeocoder)
 library(dplyr)
+library(shinythemes)
+library(ggplot2)
+library(shinyjs)
+library(zoo)
+library(tseries)
+library(forecast)
 
+### Importing Data ###
 country_df <- read.csv('data/country_data/countries of the world.csv')
 coordinate_df <- read.csv('data/country_data/countries.csv')
+covid_df <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+# Prediction import
+covid_set <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+covid_set <- covid_set %>% mutate(date_str =  as.Date(as.POSIXct(Date, format="%m/%d/%y"))) %>% arrange(date_str)
+covid_set
+# Time series import
+covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
+covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
+covid_data <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
+
+### Cleaning and formatting Data ###
+
+# removes trailing whitespaces
 country_df[,c(1)]=trimws(country_df[,c(1)])
+
+# renaming columns to conform to the leaflet lat / long naming convention
 names(coordinate_df)[names(coordinate_df) == "name"] <- "Country"
 names(coordinate_df)[names(coordinate_df) == "latitude"] <- "lat"
 names(coordinate_df)[names(coordinate_df) == "longitude"] <- "long"
+# joining lat long dictionary to our high level country data
 mergedData <<- merge(x=country_df, y=coordinate_df, by="Country", all.x=TRUE)
+# drops rows with any NA in them
 cleanData <<- mergedData[complete.cases(mergedData), ]
-print(cleanData)
+# drops rows with any empty in them
+cleanData[!apply(cleanData == "", 1, all),]
+# get a vector of names of cols for dimension input
+display_vec <<- unique(names(cleanData))
+display_vec <<- display_vec[display_vec != c('Country', 'Region')]
+# For some reson only a staggred removal is rendering in the ui ...
+display_vec <<- display_vec[display_vec != c('country', 'lat', 'long')]
 
+# test to join covid and geo data
+names(covid_df)[names(covid_df) == "Country.Region"] <- "Country"
[…]

### 2020-04-23 MSCI718 2020W Project v3.Rmd
Added Time Series Analysis and Tyler's part

@@
 
-```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=3}
+```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.5}
 
@@
 
+```{r, include=FALSE, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.5}
+# IMPORTANT must import the data and apply these manipulations first. Otherwise the function will not work
+covid_set <- read.csv('data/covid_data/covid_19_clean_complete.csv')
+covid_set <- covid_set %>% mutate(date_str =  as.Date(as.POSIXct(Date, format="%m/%d/%y"))) %>% arrange(date_str)
+```
+
+```{r, include=FALSE, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.5}
+# plotting deaths function
+# USAGE: plot_deaths('Canada')
+plot_deaths <- function(country) {
+  country_set <- covid_set %>% filter(Country.Region == country)
+  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
+  deaths <- ggplot(country_set.agg, mapping = aes(date_str, Deaths)) + geom_line() + labs(title=paste('Deaths', country), x='Date', y='Deaths')
+  return(deaths)
+}
+# plotting confirmed function
+# USAGE: plot_confirmed('Canada')
+plot_confirmed <- function(country) {
+  country_set <- covid_set %>% filter(Country.Region == country)
+  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
+  confirmed <- ggplot(country_set.agg, mapping = aes(date_str, Confirmed)) + geom_line() + labs(title=paste('Confirmed Cases', country), x='Date', y='Cases')
+  return(confirmed)
+}
+# plotting recovered function
+# USAGE: plot_confirmed('Canada')
+plot_recovered <- function(country) {
+  country_set <- covid_set %>% filter(Country.Region == country)
+  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
+  recovered <- ggplot(country_set.agg, mapping = aes(date_str, Recovered)) + geom_line() + labs(title=paste('Recovered ', country), x='Date', y='Recovered')
+  return(recovered)
+}
+plot_deaths('Canada')
+plot_confirmed('Canada')
+plot_recovered('Canada')
+```
+
+```{r, include=FALSE, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.5}
+# plotting change in deaths
+# USAGE plot_change_deaths('Canada')
+plot_change_deaths <- function(country) {
+  country_set <- covid_set %>% filter(Country.Region == country)
+  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
+  country_set.new_by_week <- country_set.agg %>% mutate(new_deaths_by_week=Deaths - lag(Deaths, 7))
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Deaths, new_deaths_by_week)) + geom_line() + labs(title=paste('Trajectory of Deaths', country), y='New Deaths (past week)', x='Total Deaths')
+  return(plot)
+}
+# plotting change in cases
+# USAGE plot_change_confirmed('Canada')
+plot_change_confirmed <- function(country) {
+  country_set <- covid_set %>% filter(Country.Region == country)
+  country_set.agg <- aggregate(select(country_set, Confirmed, Deaths, Recovered), by=country_set['date_str'], sum)
+  country_set.new_by_week <- country_set.agg %>% mutate(new_confirmed_by_week=Confirmed - lag(Confirmed, 7))
+  plot <- ggplot(country_set.new_by_week, mapping = aes(Confirmed, new_confirmed_by_week)) + geom_line() + labs(title=paste('Trajectory of Confirmed Cases', country), y='New Confirmed Cases (past week)', x='Total Confirmed Cases')
+  return(plot)
+}
+plot_change_deaths('Canada')
+plot_change_confirmed('Canada')
+```
+
 # Analysis
+## Are We Flattening the Curve?
+
+Governments all over the world have enacted social distancing measures in attempts to "flatten the curve".
+In more mathematical terms, flattening the curve means slowing the growth rate of the virus such that it is no longer on an exponential trajectory.
+This prevents the health care system from being overwhelmed and increase everyone's chances of surviving the virus.
+When we plot the confirmed cases over time it's difficult to visually tell if the growth is still exponential or if social distancing has slowed the growth.
+
+```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height=2.5}
+p1 <- plot_confirmed('South Korea') + theme(plot.title = element_text(size = 8), axis.title = element_text(size=7), axis.text = element_text(size=5)) 
+p2 <- plot_change_confirmed('South Korea') + theme(plot.title = element_text(size = 8), axis.title = element_text(size=7), axis.text = element_text(size=5)) + ggtitle("Trajectory of Confirmed Cases SK")
+y <- c(0,2,4,8,16,32,64,128,256)
+x <- c(0,2,6,14,30,62,126,254,510)
+p3 <- ggplot(data.frame(x,y), aes(x, y)) + geom_line() + geom_point() + labs(title='y=2^x', x='Total Confirmed Cases', y='New Cases (past week)') + theme(plot.title = element_text(size = 8), axis.title = element_text(size=7), axis.text = element_text(size=5))
+grid.arrange(p1, p2, p3, ncol=3)
+```
+
+A trajectory graph better illustrates the state of flattening the curve.
+The y axis represents the new confirmed cases in the past week, and the x axis represents the total confirmed cases.
+If the growth rate is exponential we can expect the graph to have a positive linear pattern.
+This can be demonstrated by graphing the function y=2^x.
+Once the growth has slowed and is no longer exponential, there a sharp dip in the graph.
+
+A country that has done really well in containing the virus is South Korea.
+As demonstrated by the linear growth, South Korea started with an exponential growth like every other country, but their effective social distancing and rapid testing measures is reflected in the graph's sharp dip.
+
+Using this graph provides a clear visual indication if a country has beaten the exponential growth and is on the right track towards flattening the curve.
+
+## Time Series Analysis
+Using ARIMA modeling, we wish to generate useful forecasts of what the situation could be like before and after various quarantine and social-distancing protocols were enacted, to see whether the "curve is flattening" and to what extent.
+For this report analyses we will only focus on 1 country.
+
+For the rest of the countries we will forecast the further spread given the latest data, to be displayed on our Rshiny dashboard.
+
+We understand that this is an extremely complex topic with numerous factors that determine the rate of spread, death, and recoveries for any given country.
+However, to simplify things we will carry on with the assumption that in order to combat the spread, there was a specific date in which preventative measures were enabled for the given country.
+
+Note: Our global data starts from January 22, 2020.
+However, in reality the first cases of the virus were recorded in November of 2019.
+Given our small dataset and the continual growth of COVID-19, we assume non-seasonality.
+
+```{r, include = FALSE, fig.height = 2.5}
+# READ AND AUGMENT DATA
+covid_data = read.csv(file.path(getwd(), "data", "covid_data", "covid_19_clean_complete.csv"), header = TRUE, stringsAsFactors = FALSE)
+covid_data$Date <- as.Date(covid_data$Date, format="%m/%d/%y")
+dates <- unique(covid_data$Date)
+firstDate = dates[1] # Jan 22, 2020
+confirmed_df <- aggregate(Confirmed ~ Country.Region + Date, FUN = sum, data=covid_data)
+death_df <- aggregate(Deaths ~ Country.Region + Date, FUN = sum, data=covid_data)
+recovery_df <- aggregate(Recovered ~ Country.Region + Date, FUN = sum, data=covid_data)
+partial_merge <- merge(confirmed_df, death_df, by = c("Country.Region", "Date"))
+covid_df <- merge(partial_merge, recovery_df, by = c("Country.Region", "Date"))
+countries <- unique(covid_df$Country.Region)
+countries_ts <- lapply(countries, function (x) {
+  country_ts <- covid_df %>% filter(Country.Region == x) %>% select(c(Confirmed, Deaths, Recovered))
+  ts(country_ts, start = as.Date("2020-1-22"), frequency = 365)
+})
+```
+
+```{r, include=FALSE, fig.height=2.5}
+# FUNCTION DEFINITIONS
+get_country_ts <- function(country_name){
+  idx = which(countries == country_name)[[1]]
+  countries_ts[[idx]]
+}
+augmented_dickey_fuller <- function(c_ts, series, regular, lag_order) {
+  if(regular==FALSE){
+    adf.test(diff(log(c_ts[, series])), alternative = "stationary", k = lag_order)
+  } else {
+    adf.test(c_ts[, series], alternative = "stationary", k = lag_order)
+  }
+}
+date_filtered_zoo <- function(country_name, endDate) {
+  wantedDates <- dates[between(dates, firstDate, as.Date(endDate))]
+  country_ts <- covid_df %>% filter(Country.Region == country_name, between(dates, firstDate, as.Date(endDate))) %>% select(Confirmed)
+  zoo(country_ts, wantedDates)
+}
+plot_zoo <- function(country_zoo, country_name) {
+  autoplot(country_zoo, main = paste("COVID-19 Confirmed Cases in", country_name)) + labs(x = "Date", y = "Number of People")
+}
+plot_ts <- function(country_ts, country_name) {
+  autoplot(country_ts, main = paste("COVID-19 Statistics in", country_name, "Since Jan 22")) + labs(x = "Time", y = "Number of People") + theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
+}
+```
+
+Let us begin by looking at the country where the disease was first encountered - China.
+By looking into how the disease spread and China handled the situation first, we may be able anticipate how other countries should handle the situation.
+
+```{r, echo=FALSE,fig.height=2.1}
+# China
+china_zoo <- date_filtered_zoo("China", "2020-4-12")
+plot_zoo(china_zoo, "China")
+```
+The first graph we want to look at is the number of confirmed cases in China to date.
+As we can see from our preliminary plot, there was a sharp spike in cases from January till March, after which the curve appears to be "flattened".
+
+From external research, it appears that China began its lockdown measures in Wuhan around January 29, which continued to rollout across the country [Greig].
+However, it takes a couple weeks for the effects of such measures to show, as people can be asymptomatic for 14 days.
+For simplicity, we will assume that February 12 (2 weeks after preventative measures) as the cut-off point for all of China.
+
+Let us forecast the number of potential cases that would exist in China if no preventative measures were taken.
+
+```{r, echo=FALSE, fig.height=2.5}
+before_china_zoo <- date_filtered_zoo("China", "2020-2-12")
+plot_zoo(before_china_zoo, "China")
+```
+To start with, we must first determine the stationarity of our time series with the Augmented Dickey-Fuller test.
+We can assume a k-value (lag) of *n/5 = 4.2 ~ 4* as recommended for non-seasonal series [Hyndman].
+```{r, echo=FALSE, warning = FALSE, fig.height=2.5}
+china_ts <- get_country_ts("China")
+before_china_ts <- head(china_ts, 21) # Feb 12 is 3 weeks from Jan 22 = 21 days
+augmented_dickey_fuller(before_china_ts, "Confirmed", TRUE, 4)
+```
+According to the test, it appears that the data is non-stationary given the high p-value.
+Let us remove the unequal variances and trend components, by taking the log and difference of the series, respectively.
+```{r, echo=FALSE, warning = FALSE, fig.height=2.5}
+augmented_dickey_fuller(before_china_ts, "Confirmed", FALSE, 4)
+```
+Now the p-value is very small and clearly significant. Hence, the data is now stationary.
+
+Let us plot the ACF and PACF charts to determine the *p* and *q* portions of our ARIMA model.  
+&nbsp;
+
+```{r, echo=FALSE,fig.height=2.5}
+c <- ggAcf(diff(log(before_china_ts[,"Confirmed"]))) + ggtitle("Confirmed Cases ACF Plot")
+d <- ggPacf(diff(log(before_china_ts[,"Confirmed"]))) + ggtitle("Confirmed Cases PACF Plot")
+grid.arrange(c,d,ncol=2)
+```
+It appears that the series is mostly represented by an Auto-Regressive (AR) model due to the entire PACF plot cutting off and the ACF plot only cutting off after the 2nd lag.
+We would assume an ARIMA(0,1,2) model.
+However, it is wise to experiment with different combinations to find the best ARIMA(p,d,q) values that have the lowest AIC and BIC.
+
+Rather than conduct this manually, we will use the `auto.arima()` function to calculate it automatically for us, with the `seasonal`, `stepwise`, and `approximation` parameters set to `FALSE` so that the best model can be chosen.
+```{r, echo=FALSE,fig.height=2.5}
+china_fit <- auto.arima(before_china_ts[, "Confirmed"], seasonal = FALSE, stepwise = FALSE, approximation = FALSE)
+china_fit
+```
+As we see from the output, our series follows an ARIMA(1,2,0) model (contrary to our initial estimation of ARIMA(0,1,2)).
+
+Let us now forecast the infection rate if China had not taken any preventative measures, for a month after February 12.  
+&nbsp; 
+```{r, echo=FALSE,fig.height=3.5}
+china_fore <- forecast(china_fit, h = 30, level = 90)
+plot(china_fore, ylab = "Number of People", main = "Forecasts for Confirmed Cases from Feb 12", xlab = "Time", xaxt = "n")
+```
+
+&nbsp;
+
+As seen from the line in blue, our forecasting till March 12 is approximately 100,000 infections (20,000 higher than the Confirmed cases by April 12), with 90% *prediction intervals* going upwards of 200,000 infections.
+
+Therefore, we can infer that the preventative measure to slow the spread of COVID-19 truly are "flattening the curve".
+
+## Map Visualization Dashboard
 
 # Conclusion
+
+## References
+
+Greig, Finlay.
+“How Long Has China Been on Lockdown?
+Here's When China and Wuhan Went on Lockdown for Coronavirus.” Edinburgh News, Edinburgh News, 6 Apr.
+2020, www.edinburghnews.scotsman.com/health/how-long-has-china-been-lockdown-heres-when-china-and-wuhan-went-lockdown-coronavirus-restrictions-approach-end-2514773.
+
+Hyndman, Robert. “Thoughts on the Ljung-Box Test.” Hyndsight, 24 Jan. 2014, www.robjhyndman.com/hyndsight/ljung-box-test/.

### 2020-04-23 MSCI718 2020W Project v3.Rmd
added Conclusion

@@
 
-Due to the current state of the world from the COVID-19 pandemic (as of April 22nd, 2020) and the rise of visualization tools being used to keep track of COVID-19 statistics across the world, we were interested in using concepts that were learned in MSCI 718 and new statistical techniques, more specifically time series analysis and interactive data visualiztion dashboard, to build a deeper understanding of the coronavirus.
+Due to the current state of the world from the COVID-19 pandemic (as of April 22nd, 2020) and the rise of visualization tools being used to keep track of COVID-19 statistics across the world, we were interested in using concepts that were learned in MSCI 718 and new statistical techniques, such as time series analysis and interactive data visualization dashboards, to build a deeper understanding of the coronavirus.
 
 Firstly, a Global COVID-19 Report dataset was sourced from Kaggle with 15,170 observations.
-This dataset included daily entries of cumulative number of cases of COVID-19 for 185 countries, each with 82 entries dating from January 22nd, 2020 to April 12th, 2020.
+This dataset included daily entries of cumulative number of cases of COVID-19 for 185 countries, each with 82 entries dating from January 22nd, 2020 to April 22nd, 2020.
 The following were the relevant variables from the dataset that were used for the analysis:
@@
 
+In addition to our report analysis, we have created an interactive visualization dashbord of the map of the world, through Rshiny. This dashboard contains 2 main components:
+1) For a given date, the cumulative amount of *Deaths*, *Recoveries*, and *Confirmed Cases* in magnitude is indicated by a red circle for each country.
+2) For a specific country chosen from the drop-down list, we display 4 graphs: Confirmed Cases, Trajectory of Confirmed Cases, Deaths, and a Forecast for the Future Number of Cases (indicated in blue from April 22 onwards).
+
+To run the app, please open the *app.R* file in our repo, and click "Run App" in RStudio.
+You may be required to install some dependencies.
+If you are unable to run the dashboard, you can also refer to our public Github page for a GIF illustrating the functionality of our tool: https://github.com/mattlianje/fluffy-katana
+
+Additionally, you can also refer to our Appendix for screenshots of the dashboard.
+
 # Conclusion
+
+The first 2 sections of this report, pertaining to flattening the curve and time series analysis for 2 Asian countries, lead us to conclusions that the preventative measures the countries have taken, such as enforcing quarantine measures and banning international travel, have indeed slowed the rate of the spread of the virus; essentially they are "flattening the curve".
+If careful measures had not been taken, our ARIMA model forecasted there would have been tens of thousands more cases by mid-March.
+
+For future endeavours we could explore the effect COVID-19 has on people of different demographics.
+As testing becomes more available and more data is released, we hope to fine tune our prediction models to more accurately predict the number of infections and provide valuable information for the public and the health care system.
 

### 2020-04-23 MSCI718 2020W Project v3.Rmd
Added more description to Data section and references

--- 2020-04-23 MSCI718 2020W Project v3.Rmd
@@
 This dataset included daily entries of cumulative number of cases of COVID-19 for 185 countries, each with 82 entries dating from January 22nd, 2020 to April 12th, 2020.
+However, the first case of COVID-19 was recorded in November 2019 [Bryner].
+In addition, the actual number of cases in a country may differ from the reported number.
+For this analysis, we assumed that this data was accurate.
 The following were the relevant variables from the dataset that were used for the analysis:
@@
 
-```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.8}
+```{r, echo = FALSE,message=FALSE, error = FALSE,fig.height=2.5}
 a <- ggplot(confirmed_df,aes(Confirmed)) + geom_histogram() + stat_bin(aes(y=..count.., label =..count..),geom="text",size=2,vjust=-0.3) + ggtitle("Frequency of Confirmed Cases") + ylab("Frequency")
@@
 
-A country that has done really well in containing the virus is South Korea.
+A country that has done really well in containing the virus is South Korea [Bhatia].
 As demonstrated by the linear growth, South Korea started with an exponential growth like every other country, but their effective social distancing and rapid testing measures is reflected in the graph's sharp dip.
@@
 
+Bhatia, Aatish. “Covid Trends.” aatishb.com/covidtrends/.
+
+Bryner, Jeanna. “1st Known Case of Coronavirus Traced Back to November in China.” LiveScience, Purch, 14 Mar. 2020, www.livescience.com/first-case-coronavirus-found.html.
+
 Greig, Finlay.

### 2020-04-23 MSCI718 2020W Project v3.Rmd
proofread

--- 2020-04-23 MSCI718 2020W Project v3.Rmd
@@
 In addition to our report analysis, we have created an interactive visualization dashbord of the map of the world, through Rshiny. This dashboard contains 2 main components:
+
 1) For a given date, the cumulative amount of *Deaths*, *Recoveries*, and *Confirmed Cases* in magnitude is indicated by a red circle for each country.
+
 2) For a specific country chosen from the drop-down list, we display 4 graphs: Confirmed Cases, Trajectory of Confirmed Cases, Deaths, and a Forecast for the Future Number of Cases (indicated in blue from April 22 onwards).
